---
title: "HW5_607 tidyr & dplyr"
author: "Chunhui Zhu"
date: "October 1, 2017"
output:
  html_document: default
  pdf_document: default
---


##The chart above describes arrival delays for two airlines across five destinations. Your task is to:

##(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You're encouraged to use a "wide" structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

```{r ecoh=F}
library("tidyr")
library("dplyr")
```


##(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

```{r}
#read data from .csv file

df<-data.frame(read.csv("hflight.csv",stringsAsFactors = FALSE))
df
```
```{r}
#fill in missed data in df$X

df[2,1]<-"ALASKA"
df[5,1]<-"AMWEST"
df
```

```{r}
#filter function remove the row which x.1 is empty, read in from second column

df<-filter(df,df$X!="")
df
```

```{r}
#gather number by airline and status

df2<- gather(df, key = "X.1", value = "count", Los.Angeles:Seattle)
colnames(df2)<-c("airline","status","city","number")
df2
```

```{r}
#reorder table base on status

df2[order(df2$airline,df2$status),]
```


```{r}
#split table base on the airline and status
#totally divid into four tables

st<-split(df2, with(df2, interaction(airline,status)), drop = TRUE)
st$ALASKA.delayed
st$`ALASKA.on time`
st$AMWEST.delayed
st$`AMWEST.on time`
```

##3. compare the arrival delays for the two airlines.

```{r}
#calculat delay rate for airlines
#built data frame t to store the data set

Alaska_delay_r<-st$ALASKA.delayed$number/(st$ALASKA.delayed$number+st$`ALASKA.on time`$number)
Amwest_delay_r<-st$AMWEST.delayed$number/(st$AMWEST.delayed$number+st$`AMWEST.on time`$number)

cities<-c("Los.Angeles","Phoenix","San.Diego","San.Francisco","Seattle")
t<-data.frame(cities, round(Alaska_delay_r,2), round(Amwest_delay_r,2))
t
```

```{r}
#The result shows Amwest has higher delay rate than Alaska in 5 cities
#Difference for taking Alaska to LA,Phonenix,San.Diego,San.Francisco,Seattle, 
#you will probably have lesser risk to delay. 

```

### Another thought for analysis

```{r}
#We lack of infomation about collected data like by seasons,daytime, 
#passenger group,from which cities, and etc.
#Even though the result from the above result shows Amwest airline 
#having more delay, it doesn't convince people taking Amwest airline 
#from city like NY will get more chance to delay than taking Alaska airline.
```


```{r}
#creat a airlines.csv file for t data set at "C:/Users/Ivy/Desktop/607/W5"

setwd("C:/Users/Ivy/Desktop/607/W5")
write.csv(t,"t.csv")
```


##(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions. Please include in your homework submission:
##The URL to the .Rmd file in your GitHub repository. and
##The URL for your rpubs.com web page.
 
Please check out two URLs from the blackboard. 

