---
title: "HW5_607"
author: "Chunhui Zhu"
date: "September 27, 2017"
output:
  pdf_document: default
  html_document: default
---

##The chart above describes arrival delays for two airlines across five destinations. Your task is to:

##(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You're encouraged to use a "wide" structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

```{r ecoh=F}
library("tidyr")
library("dplyr")
library("stringr")
```


##(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

```{r}
#read data from .csv file
df<-data.frame(read.csv("hflight.csv",stringsAsFactors = FALSE))
df 
```


```{r}
#filter function remove the row which x.1 is empty, read in from second column
#removed first column which contains airline's names
#use t function to transform rows to columns

df<-filter(df,df$X.1!="")
df<-df[,-1]
df1<-as.data.frame(t(df))
df1
```

```{r}
#remove the first row which contains header
#assign columns names 

df2<-df1[-1,]
colnames(df2)<-c("Alaska_On_time", "Alaska_delay","Amwest_On_time", "Amwest_delay")
df2
```


```{r}
#df2 data type becomes factor. 
#convert factor to numeric data type

str(df2)

df2$Alaska_On_time<-as.numeric(as.character(df2$Alaska_On_time))
df2$Alaska_delay<-as.numeric(as.character(df2$Alaska_delay))
df2$Amwest_On_time<-as.numeric(as.character(df2$Amwest_On_time))
df2$Amwest_delay<-as.numeric(as.character(df2$Amwest_delay))
df2
```


##3. compare the arrival delays for the two airlines.

```{r}
#find the delay rate for two airlines

df2$Alaska_delay_r<-round(df2$Alaska_delay/(df2$Alaska_On_time+df2$Alaska_delay),2)
df2$Amwest_delay_r<-round(df2$Amwest_delay/(df2$Amwest_On_time+df2$Amwest_delay),2)
df2[,5:6]
```

```{r}
#The result shows Amwest has higher delay rate than Alaska in 5 cities
#Difference for taking Alaska to LA,Phonenix,San.Diego,San.Francisco,Seattle, 
#you will probably have lesser risk to delay. 

df2$Diff_delay_r<-df2$Alaska_delay_r-df2$Amwest_delay_r
df2[,5:7]
```


```{r}
#generate df2 table in r
View(df2)
```

```{r}
#creat a airlines.csv file for df2 data set at "C:/Users/Ivy/Desktop/607/W5"

setwd("C:/Users/Ivy/Desktop/607/W5")
write.csv(df2,"airlines.csv")
```

### Another thought for analysis

```{r}
#We lack of infomation about collected data like by seasons,daytime, 
#passenger group,from which cities, and etc.
#Even though the result from the above result shows Amwest airline 
#having more delay, it doesn't convince people taking Amwest airline 
#from city like NY will get more chance to delay than taking Alaska airline.
```




##(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative descriptions of your data cleanup work, analysis, and conclusions. Please include in your homework submission:
##The URL to the .Rmd file in your GitHub repository. and
##The URL for your rpubs.com web page.
 
Please check out two URLs from the blackboard. 
